First Season MWA EoR Power Spectrum Results at Redshift 7

### 2016-06-29 ms.tex
SJT comments

\begin{abstract}
The Murchison Widefield Array (MWA) has collected hundreds of hours of Epoch of 
Reionization (EoR) data and now faces the challenge of overcoming foreground and 
systematic contamination to reduce the data to a cosmological measurement. We introduce 
several novel analysis techniques such as cable reflection calibration, hyper-resolution 
gridding kernels, diffuse foreground model subtraction, and quality control methods. Each 
change to the analysis pipeline is tested against a two dimensional power spectrum figure 
of merit to demonstrate improvement. We incorporate the new techniques into a deep 
integration of 32 hours of MWA data. This data set is used to place a systematic-limited 
upper limit on the cosmological power spectrum of $\Delta^2 \leq 2.7 \times 10^4$ 
mK$^2$ at $k=0.27$ h~Mpc$^{-1}$ and $z=7.1$, consistent with other published limits, and a 
modest improvement (factor of 1.4) over previous MWA results. While our result is 
systematic limited, we identify several strategies to overcome these systematics in future 
analysis.
\end{abstract}
\keywords{cosmology: observations - cosmology: reionization}
\maketitle

@@

Several probes are being used to investigate the EoR, but to date, a direct detection eludes 
researchers. Studies of the polarization of the CMB have placed integrated constraints on 
the timing of reionization \citep[e.g.][]{Planck:2015b, Planck:2015, WMAP9}. Meanwhile, 
observations of highly redshifted quasars have placed upper bounds on redshifts by which 
reionization is complete. For example, \citealt{Fan:2006a} showed that reionization must be 
complete by $z \approx 6$, a result which was confirmed independent of model by 
\citealt{McGreer:2015}.
Using a $z_{\mathrm{end}}>6$ prior, the most recent analysis of \emph{Planck} data 
found a reionization redshift $z_{\mathrm{re}} = 8.8 \pm 0.9$ for a redshift-symmetric model
($z_{\mathrm{re}} = 8.5 \pm 0.9$ for redshift-asymmetric), and a duration of $\Delta z < 2.8$
\citep{Planck:2016}.
Deep optical and infrared galaxy surveys are also beginning to 
reach the redshifts necessary to further constrain reionization \citep[e.g.][]{Bouwens:2014}, 
and the James Webb Space Telescope will improve on their sensitivity 
\citep{Gardner:2006}.

The 21 cm hyperfine transition from neutral hydrogen residing in the IGM during reionization 
offers another promising method to study the EoR. Not only is the 21cm signal a 
\emph{direct} probe of the IGM, but, due to the narrow width of the transition and the 
relationship between observed frequency and line-of-sight distance, it can be used to map 
the full three dimensional space of the epoch. (For reviews, see \citealt{Furlanetto:2006}, 
\citealt{Morales:2010}, \citealt{Zaroubi:2013}, and \citealt{Pritchard:2012}) The first generation of instruments with primary science goals 
to detect the 21cm EoR signal have been built, including the GMRT (Giant Metrewave 
Radio Telescope, \citealt{Paciga:2013}), LOFAR (LOw Frequency Array\footnote{http://
www.lofar.org}, \citealt{vanHaarlem:2013, Yatawatta:2013}), PAPER (Precision Array for Probing the Epoch of 
Reionization\footnote{http://eor.berkeley.edu}, \citealt{Parsons:2010}), and the MWA 
(Murchison Widefield Array\footnote{http://www.mwatelescope.org}, \citealt{Tingay:2013, Bowman:2013}). 
Due to relatively low signal to noise, these instruments aim for a statistical measurement of 
the EoR in the form of a cosmological power spectrum. Meanwhile, the second generation 
21cm EoR machines is on its way with the Hydrogen Epoch of Reionization Array (HERA
\footnote{http://reionization.org}, \citealt{DeBoer:2016}), which will further 
refine the power spectrum measurement in early build-out stages, but will ultimately be 
capable of imaging the ionized bubbles of reionization in its later stages 
\citep{Beardsley:2015,Malloy:2013}.

Using the first generation of instruments, several upper limits have been placed on the 21 
cm EoR signal \citep{Ali:2015, Dillon:2014, Parsons:2014, Jacobs:2015, Paciga:2013}, and 
\citealt{Pober:2015} was able to place constraints on physical reionization models. 
However, much work is to be done to understand the data produced by these arrays. 
Foreground subtraction and isolation has emerged as a priority in the field. Central to most 
analysis strategies is the concept of an ``EoR window" -- the region of Fourier space where 
spectrally smooth foregrounds have been isolated from the isotropic (spherically symmetric 
in Fourier space) cosmological signal \citep{Morales:2006, Bowman:2009}. More recent 
studies have shown the existence of a foreground ``wedge", the result of instrumental mode 
mixing, throwing power from spectrally smooth foregrounds to higher Fourier modes 
\citep{Thyagarajan:2015b, Thyagarajan:2015, Trott:2012, Liu:2014a, Liu:2014b, 
Hazelton:2013, Pober:2013, Thyagarajan:2013, Vedantham:2012, Morales:2012, Datta:2010}. Despite this 
setback, the EoR window is preserved above the wedge, and the field is pushing forward with 
several analysis pipelines under active development \citep[e.g. B.~J.~Hazelton et al. 2016, in 
prep; D.~A.~Mitchell et al. 2016, in prep; ][]{Jacobs:2016, Trott:2016, 
Dillon:2013, Trott:2014}.

This paper serves two purposes: to demonstrate several new analysis techniques and their 
impact on power spectrum estimation, and to present the first deep integration 
power spectrum from the MWA. Using a three hour test set of data, we introduce several 
novel techniques including calibration of cable reflection contamination, high resolution 
gridding kernels, subtraction of a diffuse foreground model, and development of quality 
control methods which will be crucial for deeper integrations. We apply these new 
techniques to a deep integration of data from the first semester of MWA observations 
(August, 2013 -- November, 2013) on a single EoR field and redshift range, $6.2<z<7.5$. 
While not expecting to detect the EoR with less than a hundred hours of observations, this 
intermediate integration will serve to identify and diagnose systematics, allowing for 
improved analysis in future work.

The remainder of this paper is organized as follows. In Section~\ref{sec:MWA} we briefly 
describe the MWA instrument and the observations used in this work, in Section~
\ref{sec:analysis} we describe our analysis pipeline, in Section~\ref{sec:techniques} we 
describe several novel techniques in our analysis, in Section~\ref{sec:deep} we discuss our 
efforts to apply the analysis pipeline to a deep data integration and compare with an alternative analysis pipeline for robust estimation, and we discuss future work 
in Section~\ref{sec:discussion}. Throughout this paper we use a $\Lambda$CDM 
cosmology with $\Omega_m=0.73$, $\Omega_\Lambda=0.27$, and $h = 0.7$, consistent 
with WMAP seven year results \citep{Komatsu:2011}. All distances and wavenumbers are in 
comoving coordinates.

\section{The Murchison Widefield Array and Observations}\label{sec:MWA}
The Murchison Widefield Array is one of several first generation radio interferometers with 
a primary science goal of detecting the 21cm EoR power spectrum. The remote Australian 
outback offers relative isolation from man-made radio frequency interference (RFI) such as 
FM radio, or TV stations. A recent study of the RFI environment at the Murchison Radio 
Observatory can be found in \citealt{Offringa:2015}.

@@

Calibration is accomplished within the FHD package in two steps. First we match the raw 
data to model visibilities formed using a realistic simulation of the array response to our 
foreground model. This step results in independent complex gains for each antenna, 
frequency channel, and polarization. In the second step we impose restrictions on these 
gains motivated by our understanding of the spectral response of the instrument. By 
reducing the number of free parameters in this step we increase our signal to noise on 
calibration solutions and avoid absorbing unmodeled confusion source flux into our gain 
solutions which has been shown to contaminate the EoR window \citep{Barry:2016}.

@@

The \eppsilon package performs a discrete Fourier transform (DFT) on each $r_{||}$ slice of 
the integrated HEALPix cube, forming a cube in $(\mathbf{k}_{\perp},r_{||})$, where $
\mathbf{k}_{\perp}$ represents the cosmological wavenumber in the plane perpendicular to 
our line of sight, and $r_{||}$ is the distance to the observed redshift slice along the line of 
sight. The Fourier transform along the $r_{||}$ dimension is treated separately due to 
incomplete $(u,v)$ sampling and flagged frequency channels, which leads to structure in the 
frequency sampling along any given $k_{\perp}$ pixel. We thus adopt the Lomb-Scargle 
least-squares method to determine the orthogonal eigenfunctions given our sampling 
function and estimate the total power in each $k_{||}$ mode \citep{Scargle:1982}. In 
addition, because the spectrally smooth foregrounds contain vastly more power than the 
expected EoR signal, we apply a Blackman-Harris window function prior to the $r_{||}$ to 
$k_{||}$ transform. This has the effect of trading off lower effective bandwidth for higher 
dynamic range \citep[e.g.][]{Thyagarajan:2016, Thyagarajan:2013}.

@@

It has been shown that subtracting a foreground model strictly within the main lobe of the 
primary beam will not be sufficient to suppress the power spectrum wedge and unlock the 
EoR window \citep{Thyagarajan:2015b, Thyagarajan:2015, Pober:2016}. 
We are in the process of repeating the model building described above using additional
MWA observations pointed away from our field of interest.
Until that work is complete we supplement the extent of 
our point source model using additional catalogs. We accomplish this through a hierarchical 
catalog pulling from our early KGS catalog, the MWA Commissioning Survey\footnote{The 
more complete GLEAM Survey \citep[][Hurley-Walker et al., in collaboration review]
{Wayth:2015} was not available at the time of analysis.} \citep[MWACS, ][]
{Hurley-Walker:2014}, the Culgoora catalog \citep{Slee:1977}, and the Molonglo Reference Catalog 
\citep[MRC, ][]{Large:1981}. Source fluxes are prioritized in this order based on our 
confidence in their predicted flux at 182 MHz. We first cluster the source lists to avoid 
redundant sources, using a 3.5 arcmin neighborhood radius, then select the flux from the 
highest priority catalog. Spectral indices were obtained from the MWACS and Culgoora 
catalogs when available, otherwise a two-point spectral index was estimated for Culgoora-
MRC matches or MRC-SUMSS matches. All other sources were given a spectral index of 
$-0.8$. The spectral index was used to extrapolate the flux from the catalog frequency to 
182 MHz, but a uniform spectral index of $-0.8$ is used within our band when forming 
model visibilities for calibration and foreground subtraction.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{master_catalog.pdf}
\caption{
Hierarchical catalog used for foreground subtraction. This catalog combines the source lists 
from the KGS catalog, the MWACS catalog, Culgoora sources, and the MRC, 
prioritizing in that order. The EoR0 field corresponds to the red KGS patch, while we use the 
other catalogs to fill in the sidelobes of the MWA. The size of each dot is proportional to the 
182~MHz flux of the source, clipped at 20~Jy.
\label{fig:master_catalog}
}
\end{center}
\end{figure*}

@@

This first iteration of modeling the diffuse structure within our primary field was done by 
simply using output point-source subtracted residual images of FHD from the three hour 
golden data set. We combine the images from three hours to leverage the rotation of the 
earth to improve the point spread function (PSF) of the instrument. This integrating of 
images is done in a uniformly weighted frame to minimize the effect of double instrument 
convolution (once when the data is taken, again when we use the output as a model). We 
then form a pseudo Stokes I image by adding beam-weighted East-West and North-South 
polarizations. We also integrate in frequency to form a single continuum image for our 
model. By forcing our model to contain no spectral structure we mitigate the risk of 
subtracting cosmological signal.

@@

Ultimately we need model visibilities to subtract from our data. However, rather than storing 
a model for all observations (different pointings and phase centers), we find it most simple 
to treat the model as a single HEALPix image which can be used to create model visibilities 
for each independent observation. FHD treats each pixel in the image as a point source at 
the pixel location with total flux equal to the flux density of the diffuse image times the area 
of the pixel and creates model visibilities for each snapshot in the same way that it imports 
a catalog of point sources.

@@

We demonstrate the impact of our diffuse model again by running the golden data set 
through our entire analysis pipeline with and without using the diffuse model. Because the 
foreground models used for calibration in the two runs are identical (diffuse is omitted for 
calibration), the calibration solutions were identical and therefore the dirty power spectra 
were identical as well. We compare the model and residual power spectra by direct 
subtraction in Figure~\ref{fig:diffuse_diff}. The top panel shows the model power with the 
diffuse subtracted from the model power without. The entire plot is negative (red) because 
we only added power with the diffuse model. 
Even the EoR window is red (albeit at a much lower level than the wedge) because of power
leakage from non-uniform spectral sampling and from the dynamic range limit of the 
Blackman-Harris window function.
The bottom panel is the difference in residual 
power. The wedge is completely positive (blue), indicating that the diffuse model 
successfully subtracted from the dirty visibilities. The differences in the window are positive 
and negative, indicating that 
the differences in the foreground model in this region are below the three hour noise level.

@@

We use the jackknife method -- dividing the data into many subgroups -- to filter out bad 
data and detect patterns. A powerful grouping is to divide the data into the observation day 
and ``pointing" (which direction the antennas were pointed while tracking the EoR0 field). 
We show an example of one day's worth of per-pointing power spectra in Figure~
\ref{fig:jackknife}. The pointings are labeled sequentially with $-5$ corresponding to five 
pointing steps prior to zenith transit, $0$ corresponding to zenith, and $+4$ corresponding 
to four steps after zenith. 
This numbering scheme is shown graphically in Figure~\ref{fig:pointing}.
Early in the night ($-5$ through $-3$) the bulk of the Galactic plane 
was still above the horizon, and despite being very far from our primary field of view, highly 
contaminated our observations through the sidelobes of the instrument. We then have 
relatively well-behaved pointings, with the EoR window dominated by noise, until the final 
pointing of the night when we can see evidence of the Galaxy rising again indicated by 
strong lines of power at the edge of the foreground wedge. We saw the same contamination 
on all days of observation and decided to cut all $-5$, $-4$, $-3$, and $+4$ pointings from 
our data set. In principle it may be possible to model the Galactic disk well enough to 
account for its presence, but with a lack of sufficient Galactic model, beam model, and the 
analysis tools to include it, we leave this to future work.

@@

Our final data cut was made when inspecting the residual continuum images output from 
FHD. We test for residual flux at source locations by calculating a fractional residual flux RMS 
in the following way. For each snapshot 
continuum image, we find the residual flux in the pixels of all sources greater than 0.5 Jy 
subtracted within half-beam power and scale by the integrated flux of the sources. This 
gives us a list of fractional residual fluxes for the snapshot, which we then use to calculate 
an RMS. We found that most snapshots had residual flux RMS $<10\%$ with some outliers, 
which we cut from the data. This cut largely overlapped with previous cuts, but removed an 
additional 95 observations. The cause of the high residual flux RMS is yet unknown and is 
not localized by observation day, LST, pointing, or any other jackknife we have performed.

After the cuts described above, we are left with 1,029 snapshots, or just over 32 hours of 
data. This is an aggressive cut, with the aim of removing all poor quality data and retaining 
a clean data set. As mentioned earlier, more sophisticated analysis could allow us to 
recover data cut from this study in the future (e.g. modeling and removing the Galaxy in 
observations pointing far off zenith).

@@

The 32 hour integrated residual power spectra for our three sub-bands and both 
instrumental polarizations are shown in Figure~\ref{fig:2d_ps_subband}. A number of 
features can be seen in these spectra. In all cases the foreground wedge is prominent, with 
extra power at large scales (low $k_{\perp}$). This is indication that our diffuse model can 
improve for deeper subtraction.

@@

Finally, we place our best upper limits in context with the 21cm EoR field as a whole. A 
direct comparison between measurements from different instruments is difficult due to the 
varying methods, redshifts, and scales probed. However, if we approximate the power 
spectrum in units of mK$^2$ to be roughly flat over the scales probed by the MWA, PAPER, 
and GMRT, we can glean a view of the current state of the field. The results from this 
analysis are a modest improvement over the previous best MWA results \citep{Dillon:2015}, 
which is not surprising given that this integration contains about ten times more data, but 
has a strong signal to noise on the power spectrum. Our results are also consistent with the 
GMRT limit of $\Delta^2\le6.15\times10^4$~mK$^2$ at $z=8.6$ \citep{Paciga:2013}, though 
probe significantly different periods of the EoR. The frontrunner in terms of an upper limit is 
the PAPER experiment \citep{Ali:2015, Jacobs:2015, Parsons:2014}. The PAPER-64 limit of 
$\Delta^2\le502$~mK$^2$ at $z=8.4$ is about one and a half orders of magnitude lower 
than our best limit in power units, though again probing a significantly different redshift. 
While the MWA analysis is not yet at that level of maturity, there is room for rapid 
improvement. The red lines in Figure~\ref{fig:1d_ps} indicate the noise level possible 
to reach with this data set if systematics can be overcome.

>>>>

\begin{abstract}
The Murchison Widefield Array (MWA) has collected hundreds of hours of Epoch of 
Reionization (EoR) data and now faces the challenge of overcoming foreground and 
systematic contamination to reduce the data to a cosmological measurement. We introduce 
several novel analysis techniques such as cable reflection calibration, hyper-resolution 
gridding kernels, diffuse foreground model subtraction, and quality control methods. Each 
change to the analysis pipeline is tested against a two dimensional power spectrum figure 
of merit to demonstrate improvement. We incorporate the new techniques into a deep 
integration of 32 hours of MWA data. This data set is used to place a systematic-limited 
upper limit on the cosmological power spectrum of $\Delta^2 \leq 2.7 \times 10^4$ 
mK$^2$ at $k=0.27$ h~Mpc$^{-1}$ and $z=7.1$, consistent with other published limits, and a 
modest improvement (factor of 1.4) over previous MWA results. 
From this deep analysis we have identified a list of improvements to be made to our EoR
data analysis strategies. These improvements will be implemented in the future and detailed
in upcoming publications.
\end{abstract}
\keywords{cosmology: observations - cosmology: reionization}
\maketitle

@@

Several probes are being used to investigate the EoR. 
Studies of the polarization of the CMB have placed integrated constraints on 
the timing of reionization \citep[e.g.][]{Planck:2015b, Planck:2015, WMAP9}. Meanwhile, 
observations of highly redshifted quasars have placed upper bounds on redshifts by which 
reionization is complete. For example, \citealt{Fan:2006a} showed that reionization must be 
complete by $z \approx 6$, a result which was confirmed independent of model by 
\citealt{McGreer:2015}.
Using a $z_{\mathrm{end}}>6$ prior, the most recent analysis of \emph{Planck} data 
found a reionization redshift $z_{\mathrm{re}} = 8.8 \pm 0.9$ for a redshift-symmetric model
($z_{\mathrm{re}} = 8.5 \pm 0.9$ for redshift-asymmetric), and a duration of $\Delta z < 2.8$
\citep{Planck:2016}.
Deep optical and infrared galaxy surveys are also beginning to 
reach the redshifts necessary to further constrain reionization \citep[e.g.][]{Bouwens:2014}, 
and the James Webb Space Telescope will improve on their sensitivity 
\citep{Gardner:2006}.

The 21 cm hyperfine transition from neutral hydrogen residing in the IGM during reionization 
offers another promising method to study the EoR. Not only is the 21cm signal a 
\emph{direct} probe of the IGM, but, due to the narrow width of the transition and the 
relationship between observed frequency and line-of-sight distance, it can be used to map 
the full three dimensional space of the epoch. (For reviews, see \citealt{Furlanetto:2006}, 
\citealt{Morales:2010}, \citealt{Zaroubi:2013}, and \citealt{Pritchard:2012}) The first generation of instruments with primary science goals 
to detect the highly redshift 21cm signal have been built, including the GMRT (Giant Metrewave 
Radio Telescope, \citealt{Paciga:2013}), LOFAR (LOw Frequency Array\footnote{http://
www.lofar.org}, \citealt{vanHaarlem:2013, Yatawatta:2013}), PAPER (Precision Array for Probing the Epoch of 
Reionization\footnote{http://eor.berkeley.edu}, \citealt{Parsons:2010}), and the MWA 
(Murchison Widefield Array\footnote{http://www.mwatelescope.org}, \citealt{Tingay:2013, Bowman:2013}). 
Due to relatively low signal to noise, these instruments aim for a statistical measurement of 
the EoR in the form of a cosmological power spectrum. Meanwhile, the second generation 
21cm EoR machines is on its way with the Hydrogen Epoch of Reionization Array (HERA
\footnote{http://reionization.org}, \citealt{DeBoer:2016}), 
and the low frequency Square Kilometer Array (SKA1-Low \citealt{Mellema:2013}),
which will further 
refine the power spectrum measurement in early build-out stages, but will ultimately be 
capable of imaging the ionized bubbles of reionization in its later stages 
\citep{Beardsley:2015,Malloy:2013}.

Using the first generation of instruments, several upper limits have been placed on the 21 
cm EoR signal \citep{Ali:2015, Dillon:2014, Parsons:2014, Jacobs:2015, Paciga:2013}, and 
\citealt{Pober:2015} was able to place constraints on physical reionization models. 
However, much work is to be done to understand the data produced by these arrays. 
Foreground subtraction and isolation has emerged as a priority in the field. Central to most 
analysis strategies is the concept of an ``EoR window" -- the region of Fourier space where 
spectrally smooth foregrounds have been isolated from the isotropic (spherically symmetric 
in Fourier space) cosmological signal \citep{Morales:2006, Bowman:2009}. More recent 
studies have shown the existence of a foreground ``wedge", the result of instrumental mode 
mixing, throwing power from spectrally smooth foregrounds to higher Fourier modes 
\citep{Thyagarajan:2015b, Thyagarajan:2015, Trott:2012, Liu:2014a, Liu:2014b, 
Hazelton:2013, Pober:2013, Thyagarajan:2013, Vedantham:2012, Morales:2012, Datta:2010}. Nevertheless, 
the EoR window is preserved above the wedge, and 
several analysis pipelines are under active development \citep[e.g. B.~J.~Hazelton et al. 2016, in 
prep; D.~A.~Mitchell et al. 2016, in prep; ][]{Jacobs:2016, Trott:2016, 
Dillon:2013, Trott:2014}.

This paper serves two purposes: to demonstrate several new analysis techniques and their 
impact on power spectrum estimation, and to present the first deep integration 
power spectrum from the MWA. Using a three hour test set of data, we introduce several 
novel techniques including calibration of cable reflection contamination, high resolution 
gridding kernels, subtraction of a diffuse foreground model, and development of quality 
control methods which will be crucial for deeper integrations. We apply these new 
techniques to a deep integration of data from the first semester of MWA observations 
(August, 2013 -- November, 2013), with a total of 32 hours on a single EoR field and redshift range, $6.2<z<7.5$.

Our best result is a limit on the cosmological power spectrum of $\Delta^2 \leq 2.7 \times 10^4$ 
mK$^2$ at $k=0.27$ h~Mpc$^{-1}$ and $z=7.1$. This is lower than the GMRT 40 hour
limit ($\Delta^2 \le 6.15 \times 10^4$ mK$^2$ at $z=8.6$; \citealt{Paciga:2013}), but 
signficantly higher than the latest PAPER result which 
integrated nearly 1,000 hours ($\Delta^2 \le 502$ mK$^2$ at $z=8.4$; \citealt{Ali:2015}).
Several EoR experiments are reaching sensitivity levels where very low systematics
become dominant. The PAPER team saw evidence that antenna cross-talk and foreground
leakage was limiting their result at several scales. The LOFAR team has seen excess noise
and diffuse foreground suppression due to calibration \citep{Patil:2016}. It is essential
for EoR experiments to understand and overcome these systematics in order to reach
a cosmological detection.

While not expecting to detect the EoR with less than a hundred hours of observations, our 
intermediate integration will serve to identify and diagnose systematics, allowing for 
improved analysis in future work. We list several directions to improve our modeling of
foregrounds and the instrumental response, which in turn will improve our calibration
and foreground subtraction. Our strategy to perform periodic deep integrations allows us
to continue uncovering systematics, refine algorithms and data analysis, and understand
the subtleties of the instrument to execute this challenging experiment.

The remainder of this paper is organized as follows. In Section~\ref{sec:MWA} we briefly 
describe the MWA instrument and the observations used in this work, in Section~
\ref{sec:analysis} we describe our analysis pipeline, in Section~\ref{sec:techniques} we 
describe several novel techniques in our analysis, in Section~\ref{sec:deep} we discuss our 
efforts to apply the analysis pipeline to a deep data integration and compare with an alternative analysis pipeline for robust estimation, and we discuss future work 
in Section~\ref{sec:discussion}, including a summary of planned improvements
to the analysis. Throughout this paper we use a $\Lambda$CDM 
cosmology with $\Omega_m=0.73$, $\Omega_\Lambda=0.27$, and $h = 0.7$, consistent 
with WMAP seven year results \citep{Komatsu:2011}. All distances and wavenumbers are in 
comoving coordinates.

\section{The Murchison Widefield Array and Observations}\label{sec:MWA}
The Murchison Widefield Array is one of several first generation radio interferometers with 
a primary science goal of detecting the 21cm EoR power spectrum. The remote Australian 
outback offers relative isolation from human-made radio frequency interference (RFI) such as 
FM radio, or TV stations. A recent study of the RFI environment at the Murchison Radio 
Observatory can be found in \citealt{Offringa:2015}.

@@

Calibration is accomplished within the FHD package in two steps. First we match the raw 
data to model visibilities formed using a realistic simulation of the array response to our 
foreground model. This step results in independent complex gains for each antenna, 
frequency channel, and polarization. In the second step we impose restrictions on these 
gains motivated by our understanding of the spectral response of the instrument. By 
reducing the number of free parameters in this step we increase our signal to noise on 
calibration solutions and avoid absorbing unmodeled confusion source flux density into our gain 
solutions which has been shown to contaminate the EoR window \citep{Barry:2016}. 
Recently, \citealt{Patil:2016} demonstrated excess noise and a suppression of diffuse
foregrounds in LOFAR images after calibration and subtraction of bright discrete
foreground sources. They suggested a multi-frequency calibration solution as a potential
solution to these systematics, similar to the scheme implemented here. Indeed we have
not seen evidence of these systematics through simulations and direct propagation of
our noise \citep[Hazelton et al. 2016, in prep.]{Barry:2016}.

@@

The \eppsilon package performs a discrete Fourier transform (DFT) on each $r_{||}$ slice of 
the integrated HEALPix cube, forming a cube in $(\mathbf{k}_{\perp},r_{||})$, where $
\mathbf{k}_{\perp}$ represents the cosmological wavenumber in the plane perpendicular to 
our line of sight, and $r_{||}$ is the distance to the observed redshift slice along the line of 
sight. The Fourier transform along the $r_{||}$ dimension is treated separately due to 
incomplete $(u,v)$ sampling and flagged frequency channels, which leads to structure in the 
frequency sampling along any given $k_{\perp}$ pixel. We thus adopt the Lomb-Scargle 
least-squares method to determine the orthogonal eigenfunctions given our sampling 
function and estimate the total power in each $k_{||}$ mode \citep{Scargle:1982}. In 
addition, because the spectrally smooth foregrounds contain vastly more power than the 
expected EoR signal, we apply a Blackman-Harris window function prior to the $r_{||}$ to 
$k_{||}$ transform. This has the effect of trading lower effective bandwidth for higher 
dynamic range \citep[e.g.][]{Thyagarajan:2016, Thyagarajan:2013}.

@@

It has been shown that subtracting a foreground model strictly within the main lobe of the 
primary beam will not be sufficient to suppress the power spectrum wedge and unlock the 
EoR window \citep{Thyagarajan:2015b, Thyagarajan:2015, Pober:2016}. 
We are in the process of repeating the model building described above using additional
MWA observations pointed away from our field of interest.
Until that work is complete we supplement the extent of 
our point source model using additional catalogs. We accomplish this through a hierarchical 
catalog pulling from our early KGS catalog, the MWA Commissioning Survey\footnote{The 
more complete GLEAM Survey \citep[][Hurley-Walker et al., in collaboration review]
{Wayth:2015} was not available at the time of analysis.} \citep[MWACS, ][]
{Hurley-Walker:2014}, the Culgoora catalog \citep{Slee:1977}, and the Molonglo Reference Catalog 
\citep[MRC, ][]{Large:1981}. Source flux densities are prioritized in this order based on our 
confidence in their predicted flux density at 182 MHz. We first cluster the source lists to avoid 
redundant sources, using a 3.5 arcmin neighborhood radius, then select the flux density from the 
highest priority catalog. Spectral indices were obtained from the MWACS and Culgoora 
catalogs when available, otherwise a two-point spectral index was estimated for Culgoora-
MRC matches or MRC-SUMSS matches. All other sources were given a spectral index of 
$-0.8$. The spectral index was used to extrapolate the flux density from the catalog frequency to 
182 MHz, but a uniform spectral index of $-0.8$ is used within our band when forming 
model visibilities for calibration and foreground subtraction.

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{master_catalog.pdf}
\caption{
Hierarchical catalog used for foreground subtraction. This catalog combines the source lists 
from the KGS catalog, the MWACS catalog, Culgoora sources, and the MRC, 
prioritizing in that order. The EoR0 field corresponds to the red KGS patch, while we use the 
other catalogs to fill in the sidelobes of the MWA. The size of each dot is proportional to the 
182~MHz flux density of the source, clipped at 20~Jy.
\label{fig:master_catalog}
}
\end{center}
\end{figure*}

@@

This first iteration of modeling the diffuse structure within our primary field was done by 
simply using output point-source subtracted residual images of FHD from the three hour 
golden data set. We combine the images from three hours to leverage the rotation of the 
earth to improve the point spread function (PSF) of the instrument. This integrating of 
images is done in a uniformly weighted frame to minimize the effect of double instrument 
convolution (once when the data are taken, again when we use the output as a model). We 
then form a pseudo Stokes I image by adding beam-weighted East-West and North-South 
polarizations. We also integrate in frequency to form a single continuum image for our 
model. By forcing our model to contain no spectral structure we mitigate the risk of 
subtracting cosmological signal.

@@

Ultimately we need model visibilities to subtract from our data. However, rather than storing 
a model for all observations (different pointings and phase centers), we find it most simple 
to treat the model as a single HEALPix image which can be used to create model visibilities 
for each independent observation. FHD treats each pixel in the image as a point source at 
the pixel location with total flux density equal to the surface brightness of the diffuse image times the area 
of the pixel and creates model visibilities for each snapshot in the same way that it imports 
a catalog of point sources.

@@

We demonstrate the impact of our diffuse model again by running the golden data set 
through our entire analysis pipeline with and without using the diffuse model. Because the 
foreground models used for calibration in the two runs are identical (diffuse is omitted for 
calibration), the calibration solutions were identical and therefore the dirty power spectra 
were identical as well. We compare the model and residual power spectra by direct 
subtraction in Figure~\ref{fig:diffuse_diff}. The top panel shows the model power with the 
diffuse subtracted from the model power without. 
Because our diffuse model added power to the full foreground model, the entire plot is
negative (red).
Even the EoR window is red (albeit at a much lower level than the wedge) because of power
leakage from non-uniform spectral sampling and from the dynamic range limit of the 
Blackman-Harris window function.
The bottom panel is the difference in residual 
power. The wedge is completely positive (blue), indicating that the diffuse model 
successfully subtracted from the dirty visibilities. The differences in the window are positive 
and negative, indicating that 
the differences in the foreground model in this region are below the three hour noise level.

@@

We use the jackknife method to filter out bad data and detect patterns. This involves dividing
the data into many subgroups and forming power spectra. The goal is to find observational
parameters which affect the quality of the data, such as the day or time of night.
A powerful grouping is to divide the data into the observation day 
and ``pointing" (which direction the antennas were pointed while tracking the EoR0 field). 
We show an example of one day's worth of per-pointing power spectra in Figure~
\ref{fig:jackknife}. The pointings are labeled sequentially with $-5$ corresponding to five 
pointing steps prior to zenith transit, $0$ corresponding to zenith, and $+4$ corresponding 
to four steps after zenith. 
This numbering scheme is shown graphically in Figure~\ref{fig:pointing}.
Early in the night ($-5$ through $-3$) the bulk of the Galactic plane 
was still above the horizon, and despite being very far from our primary field of view, highly 
contaminated our observations through the sidelobes of the instrument. We then have 
relatively well-behaved pointings, with the EoR window dominated by noise, until the final 
pointing of the night when we can see evidence of the Galaxy rising again indicated by 
strong lines of power at the edge of the foreground wedge. We saw the same contamination 
on all days of observation and decided to cut all $-5$, $-4$, $-3$, and $+4$ pointings from 
our data set. In principle it may be possible to model the Galactic disk well enough to 
account for its presence, but with a lack of sufficient Galactic model, beam model, and the 
analysis tools to include it, we leave this to future work.

@@

Our final data cut was made when inspecting the residual continuum images output from 
FHD. We test for residual brightness at source locations by calculating a fractional residual brightness RMS 
in the following way. For each snapshot 
continuum image, we find the residual brightness in the pixels of all sources greater than 0.5 Jy 
subtracted within half-beam power and scale by the integrated flux density of the sources. This 
gives us a list of fractional residual brightnesses for the snapshot, which we then use to calculate 
an RMS. We found that most snapshots had residual brightness RMS $<10\%$ with some outliers, 
which we cut from the data. This cut largely overlapped with previous cuts, but removed an 
additional 95 observations. The cause of the high residual brightness RMS is yet unknown and is 
not localized by observation day, LST, pointing, or any other jackknife we have performed.

After the cuts described above, we are left with 1,029 snapshots, or just over 32 hours of 
data. This is an aggressive cut, with the aim of removing all poor quality data and retaining 
a data set as clean as we can determine from our data quality checks. As mentioned earlier, more sophisticated analysis could allow us to 
recover data cut from this study in the future (e.g. modeling and removing the Galaxy in 
observations pointing far off zenith).

@@

The 32 hour integrated residual power spectra for our three sub-bands and both 
instrumental polarizations are shown in Figure~\ref{fig:2d_ps_subband}. A number of 
features can be seen in these spectra. In all cases the foreground wedge is prominent, with 
extra power at large scales (low $k_{\perp}$). This is an indication that our diffuse model can 
improve for deeper subtraction.

@@

Finally, we place our best upper limits in context with the 21cm EoR field as a whole. A 
direct comparison between measurements from different instruments is difficult due to the 
varying methods, redshifts, and scales probed. However, if we approximate the power 
spectrum in units of mK$^2$ to be roughly flat over the scales probed by the MWA, PAPER, 
and GMRT, we can glean a view of the current state of the field. The results from this 
analysis are a modest improvement over the previous best MWA results \citep{Dillon:2015}, 
which is not surprising given that this integration contains about ten times more data, but 
has a strong signal to noise on the power spectrum. Our results are also consistent with the 
40 hour GMRT limit of $\Delta^2\le6.15\times10^4$~mK$^2$ at $k=0.5\, h$~Mpc$^{-1}$ and $z=8.6$ \citep{Paciga:2013}, though 
probe significantly different periods of the EoR. The frontrunner in terms of an upper limit is 
the PAPER experiment \citep{Ali:2015, Jacobs:2015, Parsons:2014}. The PAPER-64 limit of 
$\Delta^2\le502$~mK$^2$ at $z=8.4$ from a 135 day observing campaign is about one and a half orders of magnitude lower 
than our best limit in power units, though again probing a significantly different redshift. 
While the MWA analysis is not yet at that level of maturity, there is room for rapid 
improvement. The red lines in Figure~\ref{fig:1d_ps} indicate the noise level possible 
to reach with these data if systematics can be overcome.

\subsection{Summary of planned analysis improvements}
Through the reduction of the first season of MWA high band EoR observations, we have
identified several opportunities to improve our analysis. Here we compile a list of future
improvements and their potential impact on the power spectrum results. Broadly speaking,
these improvements emphasize a need to better understand the foregrounds and the 
instrument, and are likely to be relevant to imaging based analyses in general.

\textbf{Extended point source model.} The hierarchical catalog used here was a definite
improvement over the MWACS catalog alone. This point source model will continue
to improve with the release of the GLEAM Survey, as well as dedicated MWA observations 
to target sources in the sidelobes of the EoR fields. Improving the foreground catalog
in our sidelobes will be crucial in controlling the foreground leakage into the EoR window
\citep{Pober:2016}. Incorporating these improved catalogs
into our analysis will improve both the foreground subtraction and the calibration.
